---
title: "Project2McKeonSnyder"
author: "Owen Snyder and Ashlee McKeon"
date: '2022-07-05'
output: github_document
params:
      dataChannel: "Lifestyle"
---

Render Function
```{r eval=FALSE}
rmarkdown::render("Project2McKeonSnyder.Rmd",
                  output_format = "github_document",
                  output_file = "README.md",
                  params = list(dataChannel = "Lifestyle"),
                  output_options = list(
                    html_preview = FALSE, toc = TRUE, toc_depth = 2, toc_float = TRUE)
)

```

```{r channel automation}
channel <- c("Business", "Entertainment", "Lifestyle", "Social Media", "Tech", "World", "Other")

output_file = pasteo0(channel, ".md")

params <- lapply(channel, FUN= function(x){list(channel=x)})

reports <- tibble(output_file, params)

apply(reports, margin =1,FUN=function(x)  
  {rmarkdown::render("Project2McKeonSnyder.Rmd",
                     output_format = "github_document",
                     output_file = x[[1]],
                     params = list[[2]])})
```


# Introduction
This is a full report of Project 2 conducted by Ashlee McKeon and Owen Snyder.

## Our Goal and Purpose

The main goal of this project is to create predictive models based on data from an [online news popularity site](https://archive.ics.uci.edu/ml/datasets/Online+News+Popularity). The data includes about two years worth of information regarding *Mashable* articles. This data is summarized by a set of 61 categorical and quantitative variables that will eventually help us predict the number of shares a *Mashable* article gets. We will only use a subset of these variables to help predict the shares variable using a variety of predictive models such as Random Forrest and Boosted Tree models. A main variable of interest will be the... 



## Variable Selection

Because of the large variety of variables this data set includes, we decided to only consider a subset of the variables to work with. However, it is important to note one specific set of variables we will be basing our analysis on. We will be starting with the *data_channel_is_** variables to see if they have any relation to the number of shares and the rest of the variables we select.  

The following our are variables of interest:  

* **shares** - Number of shares (this is our target variable!) 
* **n_tokens_title** - Number of words in the title
* **average_token_length** - Average length of the words in the content
* **global_rate_positive_words** - Rate of positive words in the content
* **global_rate_negative_words** - Rate of negative words in the content
* **num_imgs** - Number of images
* **num_videos** - Number of videos
* **num_hrefs** - Number of links

For categorical variables, we will be using the following varibles. In fact, we will be combining the days of the week into one variable called **dayOfWeek** and the method of its creation will be described as it arises.

* **weekday_is_monday** - Was the article published on a Monday?
* **weekday_is_tuesday** - Was the article published on a Tuesday?
* **weekday_is_wednesday** - Was the article published on a Wednesday?
* **weekday_is_thursday** - Was the article published on a Thursday?
* **weekday_is_friday** - Was the article published on a Friday?
* **weekday_is_saturday** - Was the article published on a Saturday?
* **weekday_is_sunday** - Was the article published on a Sunday?

Finally, the six data channels we will be working with come from these key variables. Also, note that we  have created one variable to represent each data channel named **dataChannel**. Its method of creation will also be described below as it arises.  

* **data_channel_is_lifestyle** - Is data channel 'Lifestyle'?
* **data_channel_is_entertainment** - Is data channel 'Entertainment'?
* **data_channel_is_bus** - Is data channel 'Business'?
* **data_channel_is_socmed** - Is data channel 'Social Media'?
* **data_channel_is_tech** - Is data channel 'Tech'?
* **data_channel_is_world** - Is data channel 'World'?

## Methods

Before we get started, it is important to briefly describe the methods wee used for analysis.  

* We will first read in the data set and manipulate the necessary variables.
* The data will be separated into a training and test set. This will be a 70/30 split.
* We will include some contingency tables at each setting of our categorical variables to get a better idea of the data at hand. We will also include summary statistics for quantitative variables. The data for these summaries will be used via the *training* data.
* We will include a variety of graphs to better visualize our data. These graphs will be described in a more subjective manner so that automation of each setting of this report makes more sense. We will still be using the *training* data.
* Finally, we will be using machine learning methods via Random Forrest Boosted Tree models. These models will be explained in greater detail in their respective section.

## Packages

Below is a list of necessary packages that will aid our data analysis and modeling.  

* `tidyverse` is a collection of useful packages designed for data science.
* `ggplot2` is an amazing way to create visually pleasing and informative graphics.
* `caret` is a set of functions that streamline the process of implementing machine learning methods.
* `rmarkdown` is a useful package for creating R Markdown documents in a variety of formats.
* `knitr` is a useful package to integrate computing and reporting.

```{r warning=FALSE,message=FALSE}
library(tidyverse)
library(ggplot2)
library(caret)
library(rmarkdown)
library(knitr)

```

# Read in Data

```{r warning=FALSE,message=FALSE}
setwd("~/Desktop")
newsData <- read_csv("OnlineNewsPopularity.csv")
newsData
## need to change working dir

```

# Create New Variables

Because we are working with six different channels across six different columns of data, we figured it would be easier to combine every channel into one column. That way, we will be working with one single channel variable that takes on unique values of Lifestyle, Entertainment, Business, Social Media, Tech, and World. This will make automation and analysis more clear. Note that the original data tells us that a data for either channel is 1 or 0. meaning that 1 means "yes" and 0 means "no".

```{r warning=FALSE,message=FALSE}
## use mutate() to create a new channel variable
newsData <- newsData %>% mutate(dataChannel = 
                              ifelse(data_channel_is_bus==1, "Business",
                              ifelse(data_channel_is_entertainment==1, "Entertainment",
                              ifelse(data_channel_is_lifestyle==1, "Lifestyle",
                              ifelse(data_channel_is_socmed==1, "Social Media",
                              ifelse(data_channel_is_tech==1, "Tech",
                              ifelse(data_channel_is_world==1, "World", "Other"))))))) %>%
                      mutate(dayOfWeek = 
                               ifelse(weekday_is_monday==1, "Monday",
                               ifelse(weekday_is_tuesday==1, "Tuesday",
                               ifelse(weekday_is_wednesday==1, "Wednesday",
                               ifelse(weekday_is_thursday==1, "Thursday",
                               ifelse(weekday_is_friday==1, "Friday",
                               ifelse(weekday_is_saturday==1, "Saturday",
                               ifelse(weekday_is_sunday==1, "Sunday", "NULL")))))))) %>%
                      mutate(weekdayWeekend = 
                               ifelse(weekday_is_monday==1, "Weekday",
                               ifelse(weekday_is_tuesday==1, "Weekday",
                               ifelse(weekday_is_wednesday==1, "Weekday",
                               ifelse(weekday_is_thursday==1, "Weekday",
                               ifelse(weekday_is_friday==1, "Weekday",
                               ifelse(weekday_is_saturday==1, "Weekend",
                               ifelse(weekday_is_sunday==1, "Weekend", "NULL"))))))))

## upon inspection, there are NA values present, this could be due to some articles not having
## one of these specific categories

#factor(newsData$dataChannel)

## convert to a factor for analysis??

#newsData

#summary(newsData)
```

## Filter Data by Channel

```{r}
newsData.busn <- newsData %>% filter(dataChannel=="Business")
newsData.ent <- newsData %>% filter(dataChannel=="Entertainment")
newsData.life <- newsData %>% filter(dataChannel=="Lifestyle")
newsData.socmed <- newsData %>% filter(dataChannel=="Social Media")
newsData.tech <- newsData %>% filter(dataChannel=="Tech")
newsData.wrld <- newsData %>% filter(dataChannel=="World")


channel.data <- newsData %>% filter(dataChannel == params$dataChannel)

```

# Split Data: Train/Test Set

In this section we will split our newsData into a training and test set. The training set will be used for model fitting and EDA while the test set will be used for predictions (verify). This will be a 70/30 split.  

NOTE: this will have to be re-ran due to the addition of the dayOfWeek variable. process will be the same but will to re-run all data chunks.

NOTE: NEED TO REMOVE VARIABLES!! i.e the variables that are based on new variables need to be deleted from data set.

```{r}
set.seed(558) ## set seed for reproducibility
#trainIndex <- createDataPartition(newsData$shares, p = 0.70, list = FALSE)
#newsTrain <- newsData[trainIndex, ]
#newsTest <- newsData[-trainIndex, ]

trainIndex <- createDataPartition(channel.data$shares, p = 0.70, list = FALSE)
newsTrain <- channel.data[trainIndex, ]
newsTest <- channel.data[-trainIndex, ]


```



# Numerical Summaries

This section is dedicated to creating basic summary statistics for our response variables, **shares**. Before performing more in depth analysis and modeling, it is important to get a better idea of how our data is distributed. For example, one should take note of means, median, and standard deviations. Another aspect to note is that if the median is greater than the mean then our data is likely skewed to the left. Conversely, if the median is less than the mean, the data is likely skewed to the right.  

```{r}
newsTrain %>% summarise(avgShares = mean(shares), medianShares = median(shares), sdShares = sd(shares))

```

# Contingency Tables

First, we may want to look at the frequency of all data channels in the newsData dataset. 

``` {r}
oneWayDataChannel <- table(newsData$dataChannel)
oneWayDataChannel 
```

We can see from the table that Tech and World are the most common, while Lifestyle and Social Media are the least common. This may suggest something meaningful exists here to look into more. 

Considering different data channels may be published more commonly on specific days of the week (e.g., business articles potentially having a higher publication frequency Monday through Friday compared to Saturday through Sunday). A two-way contingency table can be run to see if that theory holds up at the summary level of analysis. 

``` {r}
twoWayChannelDays <- table(newsData$dataChannel, newsData$dayOfWeek) 
twoWayChannelDays
```

It does appear from the table that publication frequency is much higher Monday through Friday compared to Saturday through Sunday, and that this pattern is stable across all data channels. 

# Plots

This section is dedicated to visualization by means of the `ggplot2` package. 
(edit the toc headings below)  

## Plot 1 -

We first want to visually examine if there is relationship between title length and the amount of times an article is shared. 

```{r}
correlation1 <- cor(newsData$n_tokens_title, newsData$shares)

scatterplotTitle <- ggplot(newsTrain, aes(x= `n_tokens_title`, y= `shares`)) +
  geom_point() +
  labs(title= "N_Tokens_Title  vs. Shares") + 
  geom_smooth(method = 'lm') +
  geom_smooth(col = "blue") +
  geom_text(x = 12, y = 175000, size = 4, label = paste0("Correlation = ", round(correlation1, 2)))
```

We can see from the scatterplot that there does not seem to be a linear relationship between title length and the amount of times an article is shared, as evidenced by the extremely weak correlation of 0.01. 

## Plot 2 -
What if now we are interested in determining if there is relationship between content length and the amount of times an article is shared. 

```{r}
correlation2 <- cor(newsData$average_token_length, newsData$shares)

scatterplotLength <- ggplot(newsTrain, aes(x= `average_token_length`, y= `shares`)) +
  geom_point() +
  labs(title= "Average_Token_Length  vs. Shares") + 
  geom_smooth(method = 'lm') +
  geom_smooth(col = "pink") +
  geom_text(x = 2, y = 175000, size = 4, label = paste0("Correlation = ", round(correlation2, 2)))
```

The scatterplot shows that similar to the previous analysis using title length and shares, there does not seem to be a linear relationship between content length and the amount of times an article is shared, as evidenced by the extremely weak correlation of -0.02. 

## Plot 3 -
We can use a bar graph to determine if any data channels are published more on weekdays then on weekends.

```{r}
barplotDay <- ggplot(data = newsTrain, aes(y= `weekdayWeekend`, fill = weekdayWeekend))
barplotDay + geom_bar() + 
  labs(title= "Publication on Weekday vs. Weekend")
```
We can see from the plot that there is a very apparent difference between  the number of articles published on weekdays vs. weekends, with articles published on weekdays being much higher. 

## Plot 4 -
This is a histogram of the shares variable. We wanted to get a better idea of how this response variable is distributed.  

```{r}
hist.shares <- ggplot(data = newsTrain, aes(x = log(shares)))
hist.shares + geom_histogram(bins = 45, fill = "lightblue", colour = 8) +
              ggtitle("Log Transformation of Shares")


```

## Plot 5 -

Next we will plot the number of images versus the number of shares. This will be done via a scatterplot because this will be able to give us a better visual of how these data are represented and if they are associated with one another.  

```{r}
sp1 <- ggplot(data = newsTrain, aes(x = shares, y = num_imgs))
sp1 + geom_point()

sp2 <- ggplot(data = newsTrain, aes(x = shares, y = num_hrefs))
sp2 + geom_point()

## BOXPLOT NUM IMGS BY DAY!
```


## Plot 6 -

Boxplots!  
```{r}
bp1 <- ggplot(data = newsTrain, aes(x=dayOfWeek, y=num_imgs, fill=dayOfWeek))
bp1 + geom_boxplot() + ggtitle("Boxplot: num_imgs per dayOfWeek")
## + geom_jitter() 
```


# Modeling
The main goal of our modeling section is to create predictive models for predicting the number of shares based on our variables of interest. We will be using two linear regression type models and two ensemble tree-based models. Once these models are fit, we will be able to identify the best of the four.  

First, it is important to describe in detail the methods behind each of these model types.

**Linear Models** -- Linear models are a method for predictive modeling that allows for evaluating the linear relationship between a continuous outcome variable of interest against one or more predictor variables. With linear models you are evaluating one model at a time, and can compare that model to other models of interest to determine best fit for the data. Linear regression and multiple linear regression are both examples of linear models. 
...

**Ensemble Models** -- Ensemble models are a method for taking multiple independent and different models and combining those models into one stronger model for prediction. This method can produce stronger prediction results when compared to linear models due to it's enhanced ability to reduce generalized error. Random forest and Boosting trees are both examples of ensemble models. 
 ...

## lm Model 1 -

**Full Linear Model** using all of our variables of interest. explain in more detail.

```{r}
## set up trainControl using 5-fold cross validation
trnCntrl <- trainControl(method = "cv", number=5)
## NewtrnCntrl for ENSEMBLE METHODS
NewtrnCntrl <- trainControl(method = "cv", number=5, repeats=3)




## now fit linear model for all of our chosen predictors
lmFit1 <- train(shares ~ n_tokens_title + average_token_length + global_rate_positive_words +
                         global_rate_negative_words + num_imgs + num_videos + num_hrefs,
                data = newsTrain,
                method = "lm",
                trControl = trnCntrl,
                preProcess = c("center", "scale"))
##pre process


## Now predict on the TEST data!
lmFit1.pred <- predict(lmFit1, newdata = newsTest)

```

## lm Model 2 -

**Linear Model** using all of the two- way interactions between our chosen predictors

```{r}
lmFit2 <- train(shares ~ (n_tokens_title + average_token_length + global_rate_positive_words +
                         global_rate_negative_words + num_imgs + num_videos + num_hrefs)^2,
                data = newsTrain,
                method = "lm",
                preProcess = c("center", "scale"),
                trControl = trnCntrl)
## Now predict on the TEST data!
lmFit2.pred <- predict(lmFit2, newdata = newsTest)
```

## Ensemble 1 -

**Random Forest Model**. This is a random forest model using all variables we have selected for analysis as predictors.

```{r}
rfFit <- train(shares ~ n_tokens_title + average_token_length + global_rate_positive_words +
                         global_rate_negative_words + num_imgs + num_videos + num_hrefs,
                  data = newsTrain,
                  method = "rf",
                  trControl = NewtrnCntrl,
                  preProcess = c("center", "scale"),
                  tuneGrid = data.frame(mtry= 1:7))

rfFit.pred <- predict(rfFit, newdata = newsTest)
```

## Ensemble 2 -
  
**Boosted Tree Model**. This is a boosted tree model using all variables we have selected for analysis as predictors.

```{r}
gbmGrid <-  expand.grid(interaction.depth = c(1,2,3,4), 
                        n.trees = c(25,50,100,150,200), 
                        shrinkage = 0.1,
                        n.minobsinnode = 10)

boostFit <- train(shares ~ n_tokens_title + average_token_length + global_rate_positive_words +
                         global_rate_negative_words + num_imgs + num_videos + num_hrefs,
                  data = newsTrain,
                  method = "gbm",
                  trControl = NewtrnCntrl,
                  preProcess = c("center", "scale"),
                  tuneGrid = gbmGrid,
                  verbose = FALSE)

boostFitPred <- predict(boostFit, newdata = newsTest)
```

# Model Comparison

There are many different metrics that can be utilized to determine which model should be used. For example, popular metrics include MAE, RMSE, AIC/BIC, and many more.  

This section will be dedicated to comparing our models via the RMSE metric only.  

We will be using the `predict()` function that we used above after fitting each fo the models, along with base R functions like `sqrt()` and `mean()` to calculate RMSE by hand.


```{r}
## Find RMSE for linear model fit 1
lmFit1.RMSE <- sqrt(mean((lmFit1.pred-newsTest$shares)^2))

## Find RMSE for linear model fit 2
lmFit2.RMSE <- sqrt(mean((lmFit2.pred-newsTest$shares)^2))

## Find RMSE for random forest model
rfFit.RMSE <- sqrt(mean((rfFit.pred-newsTest$shares)^2))

## Find RMSE for boosted tree model 
boostFit.RMSE <- sqrt(mean((boostFitPred-newsTest$shares)^2))

```

Now we can create a table of all of the RMSE values to get a better visual of how these values differ among model fits.
```{r}
## create data frame of all RMSE values 
all.RMSE <- data.frame(lmFit1.RMSE,lmFit2.RMSE,rfFit.RMSE,boostFit.RMSE)
colnames(all.RMSE) <- c("Linear Reg #1", "Linear Reg #2", "Random Forest", "Boosted Tree")
all.RMSE

## use which.min to find the element in the data frame with the lowest RMSE
minRMSE <- which.min(all.RMSE)
minRMSE
## now automate 
paste0("The model with the lowest RMSE value is the ", colnames(all.RMSE)[minRMSE] , " model")
```


